---
title: "Group Presentation Data Analysis"
author: "Group C"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package loading

We will need a variety of packages to load/wrangle the data, and to create visualisations. Let's load them now.

```{r load_packages}
library(readxl)
library(dplyr)
library(tidyr)
library(readr)
library(here)
library(echarts4r)
library(ggplot2)
library(thematic)
library(scales)
library(ggpubr)
```

## Data loading

Bring in the 'long' dataset:

```{r}
# Bring in the 'long' cleaned dataset

sg_long_2015_2021 <- readr::read_csv(here::here('source_data', 'SG Long format 2015 to 2021.csv'))
```

Bring in the '141 countries' long dataset:

```{r}
# Bring in the 'long' cleaned dataset

sg_long_141_countries <- readr::read_csv(here::here('source_data', 'SG long format 141 countries.csv'))
```

Bring in the augmented datasets (including 2020 data):

```{r}
long_format_158_countries <- readxl::read_excel("source_data/SG long format.xlsx", 
    sheet = "n = up to 158")

long_format_141_countries <- readxl::read_excel("source_data/SG long format.xlsx", 
    sheet = "n = 141")
```

## Visualisations

Let's begin work on creating some nice visualisations.

### Set up theme/colours

```{r}
thematic::thematic_on(bg = "#FFFFFF", fg = "black", accent = "#9BBF46")
```

### Overall trends in happiness

Bump chart - every country

```{r}
# Inspired by https://www.r-bloggers.com/2018/04/bump-chart/
unique_countries <- unique(long_format_141_countries$Country)

show.top.n <- 10

long_format_141_countries |>
  ggplot2::ggplot( aes(x = year, y = Rank, group = Country)) +
  ggplot2::geom_line(aes(color = Country, alpha = 1), size = 2) +
  ggplot2::geom_point(aes(color = Country, alpha = 1), size = 4) +
  ggplot2::geom_point(color = "#FFFFFF", size = 1) +
  ggplot2::scale_y_reverse(breaks = 1:show.top.n) +
  ggplot2::scale_x_continuous(breaks = 2015:2021, minor_breaks = 2015:2021, expand = c(.25, .25)) +
  ggplot2::coord_cartesian(ylim = c(1,show.top.n)) + 
  ggplot2::theme(legend.position = "none") +
  ggplot2::labs(x = "Year",
       y = "Ranking",
       title = "World Happiness Report Rankings",
       subtitle = "There was little change in the top 10 countries between 2015 and 2021.") +
  ggplot2::geom_text(data = long_format_141_countries %>% dplyr::filter(year == 2015),
            aes(label = Country, x = 2014.5) , hjust = .85, fontface = "bold", color = "#888888", size = 4) +
  ggplot2::geom_text(data = long_format_141_countries %>% dplyr::filter(year == 2021),
            aes(label = Country, x = 2021.5) , hjust = 0.15, fontface = "bold", color = "#888888", size = 4)
```

### Box plots by year

#### 2015

```{r}
long_format_141_countries |>
  dplyr::filter(year == 2015) |>
  ggplot2::ggplot( aes(x= reorder(Region, Score, FUN = median), y=Score, fill=Region)) + 
    ggplot2::geom_boxplot() +
    ggplot2::xlab("Region") +
    ggplot2::theme(legend.position="none") +
    ggplot2::xlab("") +
    ggplot2::xlab("") +
    ggplot2::ggtitle("Score by Region, 2015", subtitle="Sub-Sarahan Africa had the lowest median happiness.") +
    ggplot2::scale_x_discrete(labels = scales::label_wrap(20)) +
    ggplot2::coord_flip()
```

### Overall trends

```{r}
long_format_141_countries |>
ggplot2::ggplot( aes(x = year, y = Score, colour = Region)) + geom_smooth()
```

### Gini index correlation

```{r}
gini_data <- readr::read_csv(here::here('source_data', 'API_SI.POV.GINI_DS2_en_csv_v2_5181705.csv'), skip = 3)

# Pivot it to 'longer' by transforming the year column names into rows
gini_data_long <- gini_data |>
  tidyr::pivot_longer(cols = matches('\\d\\d\\d\\d'), names_to = "year", values_to = "gini") |>
  dplyr::rename("Country" = "Country Name")

# Convert the 'year' column in the gini data to numeric, to be compatible with the main dataset.
gini_data_long$year = as.numeric(gini_data_long$year)

long_format_141_countries_with_gini <- long_format_141_countries |>
  dplyr::left_join(gini_data_long, by=c("Country" = "Country", "year"= "year"))

# Remove entries with no gini
long_format_141_countries_with_gini <- long_format_141_countries_with_gini |> tidyr::drop_na(gini)

# This drops us down to 347 observations:
table(long_format_141_countries_with_gini$year)
# 2015, 2016, and 2018 have useful numbers of countries though.
```

Is there an overall correlation?

```{r}

long_format_141_countries_with_gini |>
  ggplot2::ggplot(mapping = aes(x = gini, y = Score, color = Region)) +
  ggplot2::geom_point() +
  ggplot2::ggtitle("Correlation of Gini index and happiness score (all years)") +
  ggplot2::xlab("Gini index value") +
  ggplot2::ylab("Happiness score")
```

Look at a single year instead, dropping the regions too. Generate a Pearson's R value.

```{r}
long_format_141_countries_with_gini |>
  dplyr::filter(year == 2018) |>
  ggplot2::ggplot(mapping = aes(x = gini, y = Score)) +
  ggplot2::geom_point() +
  ggplot2::ggtitle("Correlation of Gini index and happiness score (2018)", subtitle="There is no overall correlation.") +
  ggplot2::xlab("Gini index value") +
  ggplot2::ylab("Happiness score") +
  ggpubr::stat_cor(method = "pearson", label.x = 20, label.y = 8)
```
