---
title: "Group Presentation Data Analysis"
author: "Group C"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package loading

We will need a variety of packages to load/wrangle the data, and to create visualisations. Let's load them now.

```{r load_packages}
library(readxl)
library(dplyr)
library(tidyr)
library(readr)
library(here)
library(echarts4r)
library(ggplot2)
library(thematic)
library(scales)
```

## Data loading

Bring in the 'long' dataset:

```{r}
# Bring in the 'long' cleaned dataset

sg_long_2015_2021 <- readr::read_csv(here::here('source_data', 'SG Long format 2015 to 2021.csv'))
```

Bring in the '141 countries' long dataset:

```{r}
# Bring in the 'long' cleaned dataset

sg_long_141_countries <- readr::read_csv(here::here('source_data', 'SG long format 141 countries.csv'))
```

## Visualisations

Let's begin work on creating some nice visualisations.

### Set up theme/colours

```{r}
thematic::thematic_on(bg = "#FFFFFF", fg = "black", accent = "#9BBF46")
```

### Overall trends in happiness

Bump chart - every country

```{r}
# Inspired by https://www.r-bloggers.com/2018/04/bump-chart/
unique_countries <- unique(sg_long_2015_2021$Country)

show.top.n <- 10

sg_long_2015_2021 |>
  ggplot2::ggplot( aes(x = year, y = Rank, group = Country)) +
  ggplot2::geom_line(aes(color = Country, alpha = 1), size = 2) +
  ggplot2::geom_point(aes(color = Country, alpha = 1), size = 4) +
  ggplot2::geom_point(color = "#FFFFFF", size = 1) +
  ggplot2::scale_y_reverse(breaks = 1:show.top.n) +
  ggplot2::scale_x_continuous(breaks = 2015:2021, minor_breaks = 2015:2021, expand = c(.25, .25)) +
  ggplot2::coord_cartesian(ylim = c(1,show.top.n)) + 
  ggplot2::theme(legend.position = "none") +
  ggplot2::labs(x = "Year",
       y = "Ranking",
       title = "World Happiness Report Rankings",
       subtitle = "There was little change in the top 10 countries between 2015 and 2021.") +
  ggplot2::geom_text(data = sg_long_2015_2021 %>% dplyr::filter(year == 2015),
            aes(label = Country, x = 2014.5) , hjust = .85, fontface = "bold", color = "#888888", size = 4) +
  ggplot2::geom_text(data = sg_long_2015_2021 %>% dplyr::filter(year == 2021),
            aes(label = Country, x = 2021.5) , hjust = 0.15, fontface = "bold", color = "#888888", size = 4)
```

### Box plots by year

#### 2015

```{r}
sg_long_2015_2021 |>
  dplyr::filter(year == 2015) |>
  ggplot2::ggplot( aes(x= reorder(Region, Score, FUN = median), y=Score, fill=Region)) + 
    ggplot2::geom_boxplot() +
    ggplot2::xlab("Region") +
    ggplot2::theme(legend.position="none") +
    ggplot2::xlab("") +
    ggplot2::xlab("") +
    ggplot2::ggtitle("Score by Region, 2015", subtitle="Sub-Sarahan Africa had the lowest median happiness.") +
    ggplot2::scale_x_discrete(labels = scales::label_wrap(20)) +
    ggplot2::coord_flip()
```

### Overall trends

```{r}
sg_long_2015_2021 |>
ggplot2::ggplot( aes(x = year, y = Score, colour = Region)) + geom_smooth()
```
